Import and Tidy ADFG Supplied Location and Haul-out Data

### 2021-10-28 data_raw/adfg/adfg_tidydata.Rmd
NEW DATA FROM NSB & ADFG
new deployment data from ADFG via Justin and NSB provided by Andy and downloaded from movebank

There are three files provided by ADF&G

1. ADFG_SealDeploymentTable3.csv
2. ADFG_SealLocationData3.csv
3. ADFG_SealTimelineData3.csv

@@

Let's read this file in and take a look at the structure

```{r}
adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv')
```

@@

```{r}
adfg_deploy_cols <- readr::cols(
  DeployIDs = readr::col_character(),
  MapID = readr::col_character(),
  DateTagged = readr::col_datetime(format = "%m/%d/%Y %H:%M"),
  Species = readr::col_character(),
  Age = readr::col_character(),
  Sex = readr::col_character(),
  PrimaryTagType = readr::col_character(),
  FlipperTagType = readr::col_character()
)

adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv',
                                    col_types = adfg_deploy_cols) %>% 
  dplyr::mutate(DateTagged = lubridate::force_tz(DateTagged, 
                                                 'America/Anchorage')) %>% 
  dplyr::mutate(DateTagged = lubridate::as_date(DateTagged, 
                                                tz = 'America/Anchorage'))

@@

@@

```{r}
adfg_deployments <- adfg_deployments %>% 
  dplyr::rename(speno = DeployIDs,
                mapid = MapID,
                deploy_dt = DateTagged,
                primary_tag_type = PrimaryTagType,
                flipper_tag_type = FlipperTagType
                ) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(species = tolower(species),
                age = tolower(age),
                sex = tolower(sex)) %>% 
  tidyr::gather(tag_position, tag_family, c(primary_tag_type,flipper_tag_type)) %>% 
  dplyr::mutate(tag_position = ifelse(tag_position == "primary_tag_type",
                   "primary","flipper")) %>% 
  dplyr::filter(!is.na(tag_family)) %>% 
  dplyr::arrange(deploy_dt)

@@

```{r}
adfg_locations <- readr::read_csv('ADFG_SealLocationData3.csv')
```

@@

@@

```{r}
adfg_locations <- adfg_locations %>% 
  dplyr::rename(speno = DeployID,
                quality = LocationQuality,
                locs_dt = Date,
                tag_family = TagType) %>% 
  dplyr::filter(quality != "Z") %>% 
  dplyr::select(-MonthNumb) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::left_join(adfg_deployments, by = c("speno","tag_family")) %>% 
  mutate(species_code = case_when(
        species == "bearded" ~ "EB",
        species == "ribbon" ~ "HF",
        species == "spotted" ~ "PL"
    )) %>% 
    mutate(deployid = glue::glue("{species_code}",
                            "{lubridate::year(deploy_dt)}_{ptt}")) %>% 
    mutate(unique_day =
             glue::glue("{lubridate::year(locs_dt)}",
                  "{lubridate::yday(locs_dt)}",
                  .sep = "_")) %>% 
  select(-species_code) %>% 
  relocate(deployid,ptt) %>% 
  filter(species %in% c('bearded', 'ribbon', 'spotted')) %>%
    filter(lubridate::month(locs_dt) %in% c(3,4,5,6,7)) %>%
  sf::st_as_sf(coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(value = "EPSG:4326")

>>>>

There are six files provided by ADF&G

1. ADFG_SealDeploymentTable3.csv
2. ADFG_SealLocationData3.csv
3. ADFG_SealTimelineData3.csv
4. ADFG_SealDeploymentTable_NewRecords.csv
5. ADFG_SealLocationData_NewRecords.csv
6. ADFG_SealTimelineData_NewRecords.csv

@@

Let's read these files in and take a look at the structure

```{r}
adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv')
adfg_deployments_new <- readr::read_csv('ADFG_SealDeploymentTable_NewRecords.csv')
```

@@

```{r}
adfg_deploy_cols <- readr::cols(
  DeployIDs = readr::col_character(),
  MapID = readr::col_character(),
#  DateTagged = readr::col_datetime(format = "%m/%d/%Y %H:%M"),
  Species = readr::col_character(),
  Age = readr::col_character(),
  Sex = readr::col_character(),
  PrimaryTagType = readr::col_character(),
  FlipperTagType = readr::col_character()
)

adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv',
                                    col_types = adfg_deploy_cols) %>% 
  mutate(DateTagged = parse_datetime(DateTagged,format = "%m/%d/%Y %H:%M")) %>% 
  dplyr::mutate(DateTagged = lubridate::force_tz(DateTagged, 
                                                 'America/Anchorage')) %>% 
  dplyr::mutate(DateTagged = lubridate::as_datetime(DateTagged, 
                                                tz = 'America/Anchorage'))

@@

```{r}
glimpse(adfg_deployments_new)
```
```{r}
adfg_deploy_cols <- readr::cols(
  DeployIDs = readr::col_character(),
  Species = readr::col_character(),
  Age = readr::col_character(),
  Sex = readr::col_character(),
  FlipperTagType = readr::col_character()
)

adfg_deployments_new <- readr::read_csv('ADFG_SealDeploymentTable_NewRecords.csv',
                                    col_types = adfg_deploy_cols) %>% 
  mutate(DateTagged = parse_datetime(DateTagged,format = "%m/%d/%y")) %>% 
  dplyr::mutate(DateTagged = lubridate::force_tz(DateTagged, 
                                                 'America/Anchorage')) %>% 
  dplyr::mutate(DateTagged = lubridate::as_datetime(DateTagged, 
                                                tz = 'America/Anchorage'))

glimpse(adfg_deployments_new)
```

```{r}
adfg_deployments <- adfg_deployments %>% 
  bind_rows(adfg_deployments_new)
```

@@

```{r}
adfg_deployments <- adfg_deployments %>% 
  dplyr::rename(speno = DeployIDs,
                mapid = MapID,
                deploy_dt = DateTagged,
                primary_tag_type = PrimaryTagType,
                flipper_tag_type = FlipperTagType,
                flipper_last_signal = FlipperSatTagLastSignal
                ) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(species = tolower(species),
                age = tolower(age),
                sex = tolower(sex)) %>% 
  tidyr::gather(tag_position, tag_family, c(primary_tag_type,flipper_tag_type)) %>% 
  dplyr::mutate(tag_position = ifelse(tag_position == "primary_tag_type",
                   "primary","flipper")) %>% 
  dplyr::filter(!is.na(tag_family)) %>% 
  dplyr::arrange(deploy_dt)

@@

```{r}
adfg_locations <- readr::read_csv('ADFG_SealLocationData3.csv')
adfg_locations_new <- readr::read_csv('ADFG_SealLocationData_NewRecords.csv')
```

@@

```{r}
glimpse(adfg_locations_new)
```

```{r}
adfg_locs_cols <- readr::cols(
  DeployID = readr::col_character(),
  LocationQuality = readr::col_character(),
  Date = readr::col_double(),
  Latitude = readr::col_double(),
  Longitude = readr::col_double(),
  Ptt = readr::col_double(),
  TagType = readr::col_character(),
  SerialNum = readr::col_character()
)

adfg_locations_new <- readr::read_csv('ADFG_SealLocationData_NewRecords.csv',
                                  col_types = adfg_locs_cols) %>% 
  dplyr::mutate(Date = as.Date(Date, origin = "1899-12-30", tz ="UTC") %>% 
                  lubridate::as_datetime(),
                Ptt = as.integer(Ptt))

glimpse(adfg_locations_new)
```

```{r}
adfg_locations <- adfg_locations %>% 
  bind_rows(adfg_locations_new)
```

@@

```{r}
adfg_locations <- adfg_locations %>% 
  dplyr::rename(speno = DeployID,
                quality = LocationQuality,
                locs_dt = Date,
                tag_family = TagType) %>% 
  dplyr::filter(quality != "Z") %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::left_join(adfg_deployments, by = c("speno","tag_family")) %>% 
  mutate(species_code = case_when(
        species == "bearded" ~ "EB",
        species == "ribbon" ~ "HF",
        species == "spotted" ~ "PL"
    )) %>% 
    mutate(deployid = glue::glue("{species_code}",
                            "{lubridate::year(deploy_dt)}_{ptt}")) %>% 
    mutate(unique_day =
             glue::glue("{lubridate::year(locs_dt)}",
                  "{lubridate::yday(locs_dt)}",
                  .sep = "_")) %>% 
  select(-species_code) %>% 
  relocate(deployid,ptt) %>% 
  filter(species %in% c('bearded', 'ribbon', 'spotted')) %>%
    filter(lubridate::month(locs_dt) %in% c(3,4,5,6,7)) %>%
  sf::st_as_sf(coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(value = "EPSG:4326")
